The Weakly Nonlinear Saturation of the Magnetorotational Instability

### 2014-09-16 paper/mri_paper.tex
added setup diagram

The MRI evolves via a period of linear growth, followed by nonlinear saturation. The nonlinear evolution of the MRI is a complex process, and an understanding of the mechanism by which the MRI saturates remains elusive. Most previous works on MRI saturation have studied the problem in simulation. Indeed, only a few works have tackled MRI saturation analytically.

@@

Our domain is designed to be relevant to the design of the Princeton Plasma Physics Laboratory (PPPL) MRI experiment (\citei{Ji:2001kd}). The experimental apparatus is tall, one of many attributes designed to mitigate the meridional flows introduced by the endcaps. We use periodic vertical boundary conditions, so our set-up represents an idealized domain where endcap contamination does not exist. On the radial boundaries of our domain --  the inner and outer cylinders -- we apply no-slip, perfectly conducting boundary conditions.

@@

\begin{multline}
a \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{D}\mathbb{V}^*_{11}\right> \\
c \equiv \left< \mathbb{V}^\dagger \cdot \mathbf{N}^*_{31}\right> \\
\widetilde{c} \equiv \left< \mathbb{V}^\dagger \cdot \widetilde{\mathbf{N}}^*_{31}\right> \\
b \equiv \left< \mathbb{V}^\dagger \cdot \left(\mathcal{X}\mathbb{V}_{11}\right)^* \right> \\
h \equiv \left< \mathbb{V}^\dagger \cdot \left(\widetilde{\mathcal{L}}_2\mathbb{V}_{11} \, - \, \widetilde{\mathcal{L}}_1\mathbb{V}_{21} \right)^* \right> \\
g \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{L}_{3} \mathbb{V}_{11} \right> \\
\end{multline}

@@

\section{Discussion}

@@

One advantage of our

>>>>

The MRI evolves via a period of linear growth, followed by nonlinear saturation. The nonlinear evolution of the MRI is a complex process, and an understanding of the mechanism by which the MRI saturates remains elusive. Most previous works on MRI saturation have studied the problem in simulation. Indeed, only a few works have tackled MRI saturation analytically. We will briefly summarize a few of these analytic contributions.

@@

We analyze a magnetized Taylor Couette flow, representing a differentially rotating, conducting fluid between two cylinders which are driven at different speeds. We initialize an angular velocity profile which decreases outward from the axis of rotation, and a weak vertical magnetic field. These are the only two conditions required to excite MRI.

Our domain is designed to be relevant to the design of the Princeton Plasma Physics Laboratory (PPPL) MRI experiment (see \S\ref{discussion} for further discussion). We use periodic vertical boundary conditions. On the radial boundaries of our domain --  the inner and outer cylinders -- we apply no-slip, perfectly conducting boundary conditions.

\begin{figure}[h!]
\centering
\includegraphics[trim=5cm 0cm 0cm 0cm, scale=.35]{setup_diagram.eps}
\caption{Schematic diagram of our set-up, an axisymmetric magnetized Taylor Couette flow. We investigate a 2D slice of the X-Z (radial-vertical) plane. Our domain is the bolded black box, of width L. The radial dimension is solved with a basis of Chebyshev polynomials, and the vertical dimension is solved on a Fourier basis (see \S\ref{discussion}.) }
\end{figure}

@@

\beq
\begin{split}
a & \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{D}\mathbb{V}^*_{11}\right> \\
c & \equiv \left< \mathbb{V}^\dagger \cdot \mathbf{N}^*_{31}\right> \\
\widetilde{c} & \equiv \left< \mathbb{V}^\dagger \cdot \widetilde{\mathbf{N}}^*_{31}\right> \\
b & \equiv \left< \mathbb{V}^\dagger \cdot \left(\mathcal{X}\mathbb{V}_{11}\right)^* \right> \\
h & \equiv \left< \mathbb{V}^\dagger \cdot \left(\widetilde{\mathcal{L}}_2\mathbb{V}_{11} \, - \, \widetilde{\mathcal{L}}_1\mathbb{V}_{21} \right)^* \right> \\
g & \equiv \left< \mathbb{V}^\dagger \cdot \mathcal{L}_{3} \mathbb{V}_{11} \right> \\
\end{split}
\eeq

@@

\section{Discussion}\label{discussion}

@@

We compute the solutions to our derived equations pseudospectrally, using Dedalus. One advantage to using Dedalus for these computations is that the code can compute the evolution of the fully nonlinear MRI system on the same grid.

% Jeff's sims!!!!!
